---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
---

```{r,include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)

```

## Loading and preprocessing the data

File must be in your working directory. For this example we haven't unzip it, but you may use the csv file directly. 

```{r}
library(readr)
data=read_csv(file = "activity.zip")
str(data)

```


## What is mean total number of steps taken per day?

```{r}
library(dplyr,quietly = TRUE)

total_steps=data %>% group_by(date) %>% summarise(steps_per_day=sum(steps,na.rm = T))

hist(total_steps$steps_per_day,col = "blue",xlab = "Steps per Day",main = "Histogram")

png('./instructions_fig/first_histogram.png',width=500,height=300,units="px")
hist(total_steps$steps_per_day,col = "blue",xlab = "Steps per Day",main = "Histogram")

dev.off()

```


```{r}

meansteps=format(mean(total_steps$steps_per_day,na.rm = TRUE),digits=2)
mediansteps=format(median(total_steps$steps_per_day,na.rm = TRUE),digits=2)

```

The **mean** total number of steps taken per day **`r meansteps`**.  

The **median** total number of steps taken per day **`r mediansteps`**.  

## What is the average daily activity pattern?

```{r}
library(ggplot2)
data %>% 
  group_by(interval) %>% 
  summarise(steps=mean(steps,na.rm = TRUE)) %>%  ggplot(aes(x = interval,y=steps))+
    geom_line() + 
    geom_hline(aes(yintercept = mean(data$steps,na.rm=TRUE),linetype="Average"),color="blue",show.legend = TRUE)

maximo=data %>% 
  group_by(interval) %>% 
  summarise(steps=mean(steps,na.rm = TRUE)) %>% filter(steps==max(steps))
```
The interval with the maximum number of steps on average is **`r maximo$interval`** with **`r maximo$steps`** steps.  

```{r include=FALSE}
png('./instructions_fig/line_plot.png',width=500,height=300,units="px")
data %>% 
  group_by(interval) %>% 
  summarise(steps=mean(steps,na.rm = TRUE)) %>%  ggplot(aes(x = interval,y=steps))+
    geom_line() + 
    geom_hline(aes(yintercept = mean(data$steps,na.rm=TRUE),linetype="Average"),color="blue",show.legend = TRUE)

dev.off()

```

## Imputing missing values
```{r}
#Calculate the number of missing values
length(which(is.na(data$steps)))

#Create a dataset
average=data %>% 
  group_by(interval) %>% 
  summarise(steps=mean(steps,na.rm = TRUE))

navalues=left_join(data %>% filter(is.na(steps)),average,by="interval") %>% 
      mutate(steps=steps.y) %>% select(date,interval,steps)

data$steps[which(is.na(data$steps))]=navalues$steps

newdf=data
newtotal=newdf %>% group_by(date) %>% summarise(steps_per_day=sum(steps,na.rm = T)) 

hist(newtotal$steps_per_day,col = "blue",xlab = "Steps per Day",main = "Histogram")

newmeansteps=format(mean(newtotal$steps_per_day,na.rm = TRUE),digits=2)
newmediansteps=format(median(newtotal$steps_per_day,na.rm = TRUE),digits=2)
```
```{r include=FALSE}
png('./instructions_fig/second_histogram.png',width=500,height=300,units="px")
hist(newtotal$steps_per_day,col = "blue",xlab = "Steps per Day",main = "Histogram")

dev.off()
```

The **new mean** total number of steps taken per day **`r newmeansteps`** in contrast with the previous amount of **`r meansteps`** .  

The **new median** total number of steps taken per day **`r newmediansteps`** in contrast with the previous amount of **`r mediansteps`** .  

The mean and median are not the same using my method, which was assigning the average value of the interval to the intervals in which we had NAs. 


## Are there differences in activity patterns between weekdays and weekends?

```{r}
newdf=newdf %>% mutate(day=weekdays(date)) 

newdf$week=c("Weekday")
newdf$week[grep(x=newdf$day,pattern = "^S")]="Weekend"

newdf %>% group_by(week,interval) %>% summarize(steps=mean(steps)) %>%  ggplot(aes(interval,steps)) + geom_line(aes(color=week),show.legend = FALSE) + facet_grid(.~week)



```
It appears to be a diffence. Specially around de interval at 800 (presummably 8 am), where the person is more active during the weekday. This makes sense since you would assume he is moving to work. 

```{r include=FALSE}
newdf=newdf %>% mutate(day=weekdays(date)) 

newdf$week=c("Weekday")
newdf$week[grep(x=newdf$day,pattern = "^S")]="Weekend"

png('./instructions_fig/weekendcomparison.png',width=500,height=300,units="px")

newdf %>% group_by(week,interval) %>% summarize(steps=mean(steps)) %>%  ggplot(aes(interval,steps)) + geom_line(aes(color=week),show.legend = FALSE) + facet_grid(.~week)

dev.off()

```
